\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2018-01-16 eg6-letters/NIM/EG6_RTPC_NIM.tex
Advancing answers.
Advancing the review comments on NIM

To calibrate the gains, we compared the experimental ADCs to the energy 
deposited for each pad individually in GEANT4 by similar simulated tracks 
(using the same elastic events as for the drift paths calibration). This 
requires a very good GEANT4 simulation including drift paths and the spread of 
the charge along the path before reaching the pad, so that the simulated hits 
match the experimental ones. Moreover, the simulation has to match the DAQ 
features that can lead to cutting out some hits. After setting the simulation 
properly, we compared simulation to experiment on an event by event basis as 
shown in Figure~\ref{fig:EVENT_adc_tdc}. The gain for each pad was calculated 
as the ratio of the measured ADCs to the simulated deposited energy. Then, this 
set of the gains was applied to the experimental data. We noticed that some 
pads record lower ADCs than expected. For this reason, we wanted to make sure 
that the recorded ADCs by a given pad are similar to the ADCs recorded by other 
pads in the same track. At this step, we compared the mean ADCs of each
pad to the mean ADCs of the whole track. This ratio is collected for each pad 
from a sample of elastic events and a gain correction factor is extracted and 
applied to the first extracted gain ratio giving a final gain for each readout 
pad. Final results are shown in Figure~\ref{fig:dedx_p_exp_2nd}, where energy 
loss of particles is plotted against momentum over charge ratio. One can 
clearly see the band for $^4$He in its expected position.

\begin{figure}[!t]
   \centering
\includegraphics[scale=0.42]{Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

@@

The first is an oscillatory noise located early in the readout time window, 
shown in the top panel of Figure~\ref{fig:noise} for a particularly noisy 
channel.  Its amplitude is similar to those of real tracks. About 18\% of the 
readout channels exhibit large contributions from this noise characteristic.  
Due to its unique time-energy correlation for the given channels, the noise 
could be removed on an event by event and channel by channel basis without 
significant loss of good signals. The result of the
procedure is illustrated in the bottom panel of Figure~\ref{fig:noise}.

@@

Elastic scattering was used to estimate the tracking resolution of the RTPC 
based on the residual between the expected and measured $^4$He tracks.  The RTPC
resolutions, after removing contributions from the electron, are shown in Table
~\ref{tab:reso}, and are very similar for the two halves of the RTPC.  Note that the
$\theta$- and $z$-resolutions are highly correlated.
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|cccc|}
  \hline
& $\sigma_{z}$ &  $\sigma_{\theta}$ & $\sigma_{\phi}$ & $\sigma_{p}/p$\\
\hline
Left &  5.3 mm & 3.8$^{\circ}$ & 1.9$^{\circ}$ & 9$\%$ \\
Right & 6.5 mm & 4.0$^{\circ}$ & 1.9$^{\circ}$ & 8$\%$\\
\hline
\end{tabular}
\caption{The resolutions of the two modules of the RTPC for $z$-vertex, polar and azimuthal angles, and momentum.}
\label{tab:reso}
\end{center}
\end{table}

>>>>

To calibrate the gains, we compared the experimental ADCs to the energy 
deposited for each pad individually in GEANT4 by similar simulated tracks 
(using the same elastic events as for the drift paths calibration). This 
requires a detailed simulation, including the drift paths and the diffusion of 
the charges along the path before reaching the pad to 
match the experimental data. We implemented in the GEANT4 simulation the
drift path we obtained from the calibration described above and implemented an ad 
hoc diffusion to match the average number of hits recorded in the simulation with
the one from experimental data. To perform this step properly, we implemented in
the simulation the details of the DAQ process to record hits.

We then compared this simulation to experiment on an event by event basis as 
shown in Figure~\ref{fig:EVENT_adc_tdc}. The gain for each pad was calculated 
as the ratio of the measured ADCs to the simulated deposited energy. This 
provided our initial set of gains for each pads and it was applied to the 
experimental data. However, after this calibration some pads recorded lower ADCs 
than expected. So, we added a correction factor obtained by comparing  
the energy deposit on a given pad to its neighbours. To do so, we compared, for 
each track, the energy deposit on a pad to the energy deposit in the whole track. 
The gain correction factors are extracted by averaging this ratio for a sample 
of elastic events. Final results are shown in Figure~\ref{fig:dedx_p_exp_2nd}, where energy 
loss of particles is plotted against momentum over charge ratio. One can 
clearly see the band for $^4$He in its expected position\footnote{The lower peak 
   in the left module of the RTPC comes from an unidentified problem in this half 
   of the detector that concerns 7\% of all the elastic events. 
   Our best guess is that these events are linked with a high voltage supply 
   issue, that would sometimes bring the GEM gain down. These events pass all 
   the elastic requirements and we found nothing that differenciates them 
   from other events but their low recorded ADCs. For 
   the calibration procedures, these events were therefore excluded.}.

\begin{figure}[!t]
   \centering
\includegraphics[scale=0.42]{Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a single track. The different colors indicate the signals from different pads, 
the same color in the top and bottom figure indicate that the signal was registered
on the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

@@

The first is an oscillatory noise located early in the readout time window, 
shown in the top panel of Figure~\ref{fig:noise} for a particularly noisy 
channel. Its amplitude is similar to those of real tracks. About 18\% of the 
readout channels exhibit large contributions from this noise.  
Due to its unique time-energy correlation, we use a patern
recognition algorithm and discard the hits coming from the channels that display 
this behavior on an event by event basis. The result of the
procedure is illustrated in the bottom panel of Figure~\ref{fig:noise}.

@@

Elastic scattering was used to estimate the tracking resolution of the RTPC 
based on the residual between the expected and measured $^4$He tracks.  The RTPC
resolutions, after removing contributions from the electron, are shown in Table
~\ref{tab:reso}, and are very similar for the two halves of the RTPC.  Note that the
$\theta$- and $z$-resolutions are highly correlated.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|cccc|}
  \hline
& $\sigma_{z}$ &  $\sigma_{\theta}$ & $\sigma_{\phi}$ & $\sigma_{p}/p$\\
\hline
Left &  5.3 mm & 3.8$^{\circ}$ & 1.9$^{\circ}$ & 9$\%$ \\
Right & 6.5 mm & 4.0$^{\circ}$ & 1.9$^{\circ}$ & 8$\%$\\
\hline
\end{tabular}
\caption{The resolutions of the two modules of the RTPC for $z$-vertex, 
polar and azimuthal angles, and momentum.}
\label{tab:reso}
\end{center}
\end{table}

### 2018-01-16 eg6-letters/NIM/review-1/review-1.tex
Advancing the review comments on NIM

\subsection*{ Chapter 4}
	/paragraph 4.1
\item I would really prefer to call it drift "velocity" instead of drift 
   "speed"\\
\textcolor{blue}{What is calibrated here is mainly a speed, as we are just looking 
how fast electrons are going without regard at there direction, which is looked into 
in the following paragraph.}

@@

        /paragraph 4.3
\item The gain calibration is very quick. I would like more details, e.g. 
   "setting the simulation properly" intends only GEANT4 or also MAGBOLTZ or 
   GARFIELD? Were they used to simulate the drift and avalanche formation of 
   the electrons? And the diffusion from the gas? Was a specific code written 
   to simulate the electronics?\\
\textcolor{blue}{Only GEANT4 simulation was used in performing the gain 
   calibration, where the drift paths are updated in the simulation. The 
   electrons avalanche formation and the gas diffusion are performed though 
GEANT4. We did not implement a detailed description to simulate the 
electronics, but we did apply similar conditions to match the real DAQ, such as 
the rejection of the signals from bad pads and conditions to ensure that each 
pad must have at least 3 consecutive time bins, with ADC values above the 
threshold, 35 ADCs, in order to be recorded. The text is cleared with more 
details. }

\item line 330: "correction factors" are mentioned. On what? Why haven't only 
   good tracks been used for calibration? Lack of statistics?\\
\textcolor{blue}{Correction factor on the extracted gain for each pad. We used 
the elastic $^4$He tracks as we could calculate the recoil $^4$He kinematics 
from the scattered electron in CLAS to avoid any discrepancy from the RTPC.}

\item When it says that the pad charge is "compared" to the neighboring ones, 
   what happens next? Is it the noise reduction step described later? If so, 
   please write it. \\
\textcolor{blue}{The gain extraction is following this order: \\
- pass 1: the gain of each pad is defined as the ratio between the mean 
recorded experimental ADCs to the mean simulated ADCs in the same track. Then 
for each pad, the gain ratios were collected from all the identified elastic 
tracks and these gains were fitted by a Landau function to give a
  first pass gain for each pad as the most probable value of the fit.\\
 - pass2: to make sure that the recorded ADCs by a given pad are similar to the 
 recorded ADCs by other pads in the same track, we refined this calibration 
 after the extracted gains were applied to the experimental data. At this step, 
 we compared the mean ADCs of each pad to the mean ADCs of the whole track. 
 This ratio is collected from all the elastic events and a gain correction 
 factor is extracted and applied giving a final gain for each pad.
  }

\item Figure 11: I don't understand it. How many tracks does the histo contain? 
   If the single pad has 128 channels, how many channels does the single bin 
   contain? Moreover, is the dE calculated on pad by pad basis? In that case, 
   why don't you put just one point in the histo for each pad? What is the mean 
   number of channels or pads fired by one track? Sorry for the list of 
   questions, but I could not understand the figure.\\
   \textcolor{blue}{- It is for only one track (written in the caption).
   - We dont understand your second question. A pad is a channel. Do you mean 
how many pads are fired for a single track? Figure \ref{npd_track} shows the 
number of active pads as a function of the momentum for the collected elastic 
$^4$He tracks.\\
- dE is calculated from the ADCs of the individual pads that read out hits 
  corresponding to the same track. One needed to know the individual step size 
  very precisely to put dEdx from the individual pads, which is not available.  
  - Last question is addressed by figure \ref{npd_track}. The mean value is 
    around 10 pads. }

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.45]{npd_p_elastic.png}
   \caption{The number of active pads versus the measured momentum of the 
   elastic $^4$He tracks. } \label{fig:npd_track}
\end{figure}

\item Figure 12 - please, put it after figure 11. Moreover, please change dEdx 
   to dE/dx on y axis. Why are the left and right pictures so different?  
   Talking about the deuteron band which is present only in the left figure and 
   the smear on the top part of the right picture. \\
\textcolor{red}{- The figures layout will be updated.\\}

\textcolor{blue}{- The y-axis label on figure 12 is updated.\\}
\textcolor{blue}{- Regarding the lower peak in the left module of the RTPC, 
   these events pass all the elastic requirements but for some reasons they 
   have lower ADC values.  They represent around 7\% of all the elastic events.  
   After extended studies, the nature of these particles is not identified yet.  
   We note that this is a global phenomenon in the left module as 94\% of the 
   left module's pads are involved in both some low and high dEdx events. For 
   the calibration procedures, the events with low dEdx were excluded as we do 
not fully understand their nature.}

@@

\item Please detail a little bit more the procedure to remove the oscillatory 
   noise. It says "event by event and channel by channel": if the ADC threshold 
   is passed then a noise level is subtracted? If so, how is it calculated?\\
\textcolor{blue}{The oscillatory noise was removed just by rejecting all hits 
from a channel if too many of its hits fall on the noise curve in a given 
event. No threshold/pedestal concept, just hit rejection if it looks like 
noise. }

/paragraph 5.2
\item How was "10.5 mm" cut chosen? is it in xy plane or in 3-dimensions?\\
\textcolor{blue}{The distance between hits are calculated in 3D. This cut has 
been chosen because it enables us to jump longer gaps from dead areas of the 
readout pads without increasing the level of spurious tracks.}

\item How was the "10 hits" cut chosen? What is the mean number of primary 
   ionization in this gas mixture?\\
   \textcolor{blue}{Figure \ref{fig:nb_hits} shows the number of hits 
   distribution for the collected elastic $^4$He tracks from the 1.2 GeV 
dataset. The mean number of hits is around 135 and the 10 hits cut at the 
reconstruction level is a loose cut to reduce the level of the spurious 
tracks.}

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.45]{nb_hits.png}
   \caption{the number of hits distribution for the collected elastic $^4$He 
   tracks from the 1.2 GeV dataset.} \label{fig:nb_hits}
\end{figure}

\item < dE/dx > is computed here as corrected Etot divided total track length. 
   Usually the mean dE/dx is computed with the truncated mean of the specific 
   energy losses of each hit. Isn't it possible to compute it like this here? 
   dx for each hit can be (maybe) computed from the difference in the hit 
   times. Sum $dE_{i}$/Sum $dx_{i}$ is not equal to Sum ($dE_{i}/dx_{i}$).\\
\textcolor{blue}{We actually did compute dEdx with  path-length and ADC- 
truncations, and a linear fit to ADC vs path-length to get slope and fit 
quality and dedx from fit. Unfortunately, it never seemed to give us much 
improvement compared to Sum $dE_{i}$/Sum $dx_{i}$. }

@@

        /paragraph 6.2
\item I understand that efficiency is a ratio between two integrals, each 
   integral being the area of the Gaussian which fits the recoil mass 
   distribution in the inclusive and exclusive case. If so, would it be 
   possible to please draw both fits on the figure 14 and provide the integrals 
   to compute the mean efficiency? If I misunderstood, please comment.\\
\textcolor{red}{Nathan, do you still have the individual integrals? }

@@

\item Is it "high rate environment" or "high readout rate"? They are connected, 
   but what is the rate the RTPC underwent during data taking?\\
\textcolor{red}{Can we find such information in the logbook?}

\item I understand the resolution and efficiency have been evaluated with the 2 
   GeV electrons. Is it foreseen to analyze also the 6 GeV electron runs, which 
   were the actual data taking? Is there an evaluation of the foreseen RTPC 
   performances in that case?\\
\textcolor{blue}{Typically, the elastic reaction is used for such performance 
evaluation. But the elastic cross sections is really small at 6 GeV beam 
energy.}

@@

>>>>

\subsection*{ Chapter 4}
	/paragraph 4.1
\item I would really prefer to call it drift "velocity" instead of drift 
   "speed"\\
\textcolor{blue}{What is calibrated here is mainly the drift time, based 
on your next question as well, we clarified the text about these.} 
\textcolor{red}{To be implemented}

@@

        /paragraph 4.3
\item The gain calibration is very quick. I would like more details, e.g. 
   "setting the simulation properly" intends only GEANT4 or also MAGBOLTZ or 
   GARFIELD? Were they used to simulate the drift and avalanche formation of 
   the electrons? And the diffusion from the gas? Was a specific code written 
   to simulate the electronics?\\
\textcolor{blue}{The section has been extended to provide more details. We 
implemented specific codes in our GEANT4 simulation to have a realistic 
description of the electron drifts and of the DAQ process.}

\item line 330: "correction factors" are mentioned. On what? Why haven't only 
   good tracks been used for calibration? Lack of statistics?\\
\textcolor{blue}{We improved the gains by using a method based only on data,
the text has been clarified on the topic. All these calibration procedures
are performed with good tracks only.}

\item When it says that the pad charge is "compared" to the neighboring ones, 
   what happens next? Is it the noise reduction step described later? If so, 
   please write it. \\
\textcolor{blue}{The gain extraction is following this order: \\
- pass 1: the gain of each pad is defined as the ratio between the mean 
recorded experimental ADCs to the mean simulated ADCs in the same track. Then 
for each pad, the gain ratios were collected from all the identified elastic 
tracks and these gains were fitted by a Landau function to give a
  first pass gain for each pad as the most probable value of the fit.\\
 - pass2: to make sure that the recorded ADCs by a given pad are similar to the 
 recorded ADCs by other pads in the same track, we refined this calibration. For this, 
 we compared the mean ADCs of each pad to the mean ADCs of the whole track. 
 This ratio is collected from all the elastic events and a gain correction 
 factor is extracted and applied giving a final gain for each pad.
  }

\item Figure 11: I don't understand it. How many tracks does the histo contain? 
   If the single pad has 128 channels, how many channels does the single bin 
   contain? Moreover, is the dE calculated on pad by pad basis? In that case, 
   why don't you put just one point in the histo for each pad? What is the mean 
   number of channels or pads fired by one track? Sorry for the list of 
   questions, but I could not understand the figure.\\
   \textcolor{blue}{This figure only displays one track. Pads are single channels, 
the bins correspond to time (100 ns per bin). The dE is calculated overall in general,
we use this pad by pad method only for calibration. The average number of pad per 
track is 9, with most of the good tracks between 6 and 14.}

\item Figure 12 - please, put it after figure 11. Moreover, please change dEdx 
   to dE/dx on y axis. Why are the left and right pictures so different?  
   Talking about the deuteron band which is present only in the left figure and 
   the smear on the top part of the right picture. \\
\textcolor{red}{The figures layout will be updated.\\}

\textcolor{blue}{The y-axis label on figure 12 is updated. Regarding the lower 
   peak in the left module of the RTPC, it is likely not deuterons. We added
   a footnote to clarify the question.}

@@

\item Please detail a little bit more the procedure to remove the oscillatory 
   noise. It says "event by event and channel by channel": if the ADC threshold 
   is passed then a noise level is subtracted? If so, how is it calculated?\\
\textcolor{blue}{The oscillatory noise was removed just by rejecting all hits 
from a channel if too many of these hits fall on the noise curve in a given 
event. We did not use a threshold or pedestal concept, just a pad rejection 
if we observe the oscillating noise in the event, the text was clarified.}

/paragraph 5.2
\item How was "10.5 mm" cut chosen? is it in xy plane or in 3-dimensions?\\
\textcolor{blue}{The distance between hits are calculated in 3D. This cut has 
been chosen because it enables to jump a one pad gap in case of a dead 
channel.}

\item How was the "10 hits" cut chosen? What is the mean number of primary 
   ionization in this gas mixture?\\
   \textcolor{blue}{This cut is set to avoid tracks made out of random noise and
to limit computing time. The number of hits in good tracks at the end of the 
analysis is always above 30, with a mean around 135, so this cut has no impact 
on the good track selection.}

\item < dE/dx > is computed here as corrected Etot divided total track length. 
   Usually the mean dE/dx is computed with the truncated mean of the specific 
   energy losses of each hit. Isn't it possible to compute it like this here? 
   dx for each hit can be (maybe) computed from the difference in the hit 
   times. Sum $dE_{i}$/Sum $dx_{i}$ is not equal to Sum ($dE_{i}/dx_{i}$).\\
\textcolor{blue}{We compute dEdx with path-length and ADC- 
truncations, and a linear fit to ADC vs path-length to get slope and fit 
quality and dedx from fit. Unfortunately, it never seemed to give us much 
improvement compared to Sum $dE_{i}$/Sum $dx_{i}$. Note that, by the nature of 
the FADC electronics used, our hits always correspond to 100 ns time bins such that
each measured ADCs are almost dE/dx measurements already.}

@@

        /paragraph 6.2
\item I understand that efficiency is a ratio between two integrals, each 
   integral being the area of the Gaussian which fits the recoil mass 
   distribution in the inclusive and exclusive case. If so, would it be 
   possible to please draw both fits on the figure 14 and provide the integrals 
   to compute the mean efficiency? If I misunderstood, please comment.\\
\textcolor{blue}{This study has been performed by a student a long time
ago and we do not have all the individual fits anymore.}

@@

\item Is it "high rate environment" or "high readout rate"? They are connected, 
   but what is the rate the RTPC underwent during data taking?\\
\textcolor{red}{The rate of protons is of MHz in the chamber with a readout of
about 3kHz, so one could argue that it is both. Our point here is to emphasis
the high rate however, as these slow protons deposit a lot of energy in the 
chamber.}

\item I understand the resolution and efficiency have been evaluated with the 2 
   GeV electrons. Is it foreseen to analyze also the 6 GeV electron runs, which 
   were the actual data taking? Is there an evaluation of the foreseen RTPC 
   performances in that case?\\
\textcolor{blue}{It is not possible to do such studies with the 6 GeV beam, because 
of the lack of an exclusive process to use as elastic cross sections drops dramatically 
in our acceptance at such high energy. We would have loved to and tried many alternate
channels, but none was detectable with reasonnable statistics.}

@@

\textcolor{blue}{We would like to thank the reviewer for the detailed remark that
helped improving the manuscript significantly.}
